---
title: "Pli drought mortality"
author: "Hardy Griesbauer"
date: "20/09/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Install package

Run the below code if you would like to re-run the following analyses using R.

```{r message=FALSE}

# Run the following code to install package 
# remotes::install_github("https://github.com/hgriesbauer/pliDrought2019")

# Load libraries
library(pliDrought2019)
library(dplyr)
library(ggplot2)
library(tidyr)

```

## Data description

Thanks Rachel for your great notes and summaries!

In July 2019, we sampled data from `r nrow(pliDat)` trees, in `r nrow(pliDat)/2` paired plots, distributed over `r length(unique(pliDat$Opening))`.  Each plot contained one dead pine tree, with drought as the likely mortality agent, and one live tree.  The following criteria were used to identify tree pairs:

1. Dead tree was determined to have been killed by drought within past two years;
2. Dead tree had <=20% live needles (in other words, 80+% of needles on tree were red/dead);
3. Live tree had >=80% live needles;
4. Both trees had a similar microtopographic position;
5. Both trees had similar diameter, height, crown class;and
6. Both trees were free of visible forest health factors (some trees with stem disease were sampled).

## Preliminary analysis

### Tree-level measurements
For each tree, we collected the following measurements:

1. Tree # - Tree #1 is always the dead tree, Tree #2 is live
2. Distance to live partner - Distance between trees in plot, m
3. Crown class - Dominant or codominant trees were sampled
4. Crown volume - This compares volume between live and dead trees within a plot.  Live tree crown was assigned a volume of 100%, and the dead tree % is relative to live crown percentage.  E.g., a dead tree crown volume % of 50% means its crown volume (when alive) would have been 1/2 of the live tree.
5. Longest branch - Ocular estimate of longest branch within crown, cm
6. B - Height to bottom of crown (in case of dead trees, where live crown was),m
7. Percentage of live needles - Percentage of live needles to dead needles in crown

```{r}

# BOXPLOTS
pliDat %>% 
  dplyr::select(`Opening`,`Plot`,`Tree #`,
                `Crown Volume (%)`,`Longest Branch (cm)`,
                `Height (m)`,`DBH (cm)`,`B (m)`,`% live needles`) %>% 
  pivot_longer(cols=4:9,names_to="var") %>% 
  mutate(`Tree status`=factor(`Tree #`,labels=c("Dead","Live"))) %>% 
  ggplot(aes(y=`value`,x=`Tree status`,fill=`Tree status`))+
  geom_boxplot()+
  facet_wrap(~var,scales="free_y")+
  theme_bw()

```

Some differences from these boxplots:
1. Height to live crown (B) looks to be higher in live trees;
2. DBH is higher in live trees (this seems intuitive, given that the trees have been growing, whereas dead trees may have stopped growing 2 years ago);
3. Live trees have longer crown branches than dead trees.

Heights seems to be relatively equal between live/dead trees.

The % live needles and % crown volume differences seem intuitive.


### Cores
We also collected a single tree core from each tree, at 1.3m off the ground.  Dr. Lisa Wood at UNBC is processing the cores in fall 2019 and will provide annual measures of:
1. Ring density (g/cm3)
2. Earlywood and Latewood density
3. Ring width (mm)
5. Earlywood and Latewood width (mm)


