---
title: "Pli drought mortality"
author: "Hardy Griesbauer"
date: "20/09/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Install package

Run the below code if R package hasn't already been installed:

```{r}

# Run the following code to install package 
# remotes::install_github("https://github.com/hgriesbauer/pliDrought2019")
library(pliDrought2019)
library(dplyr)
library(ggplot2)
library(tidyr)

```

## Data description

Thanks Rachel for your great notes and summaries!

In July 2019, we sampled data from `r nrow(pliDat)` trees, in `r nrow(pliDat)/2` paired plots, distributed over `r length(unique(pliDat$Opening))`.  Each plot contained one dead pine tree, with drought as the likely mortality agent, and one live tree.  The following criteria were used to identify tree pairs:

1. Dead tree was determined to have been killed by drought within past two years;
2. Dead tree had <=20% live needles (in other words, 80+% of needles on tree were red/dead);
3. Live tree had >=80% live needles;
4. Both trees had a similar microtopographic position;
5. Both trees had similar diameter, height, crown class;and
6. Both trees were free of visible forest health factors (some trees with stem disease were sampled).

## Preliminary analysis
```{r}

# BOXPLOTS
pliDat %>% 
  dplyr::select(`Opening`,`Plot`,`Tree #`,
                `Crown Volume (%)`,`Longest Branch (cm)`,
                `Height (m)`,`DBH (cm)`,`B (m)`,`% live needles`) %>% 
  pivot_longer(cols=4:9,names_to="var") %>% 
  mutate(`Tree status`=factor(`Tree #`,labels=c("Dead","Live"))) %>% 
  ggplot(aes(y=`value`,x=`Tree status`))+
  geom_boxplot()+
  facet_wrap(~var,scales="free_y")

```


